Vulnerability Analysis of Container Images in Neuroimaging

### 2020-05-22 paper.tex
Some comments

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip}
\newcommand{\rom}[1]{\lowercase\expandafter{\romannumeral #1\relax}}
\newcommand{\TG}[1]{\color{blue}From Tristan: #1\color{black}}

@@

@@

This study involves four image scanning experiments.
The first experiment scans the images as is to report the existing vulnerabilities
inside the images. The scan report of
these images shows a large number of existing vulnerabilities which may pose
a potential security risk. These vulnerabilities exist due to the outdated
packages that are present inside the images. So, the next step is to
take the necessary steps to improve security in the images. Therefore, there exist
two possible solutions.
The first one involves updating
all these images and then rescanning them, which constitutes the second experiment of our study. The next 
experiment involves shrinking the size of the image
meaning that removing all unnecessary packages from the image to see the effect on the number of
vulnerabilities present inside the image, which will be presented as our third experiment.
Lastly, our fourth experiment involves both updating and shrinking the image size and then
rescanning for the vulnerabilities.

@@

@@

%\begin{comment}
%Singularity offers mobility of compute by facilitating portable environments 
%through a single image file~\cite{kurtzer2017singularity}. Once Singularity image
%is created, it is hashed using SHA-256. Hence, it cannot be changed
%once it is created, consequently, it can be used for reproducing the results of
%scientific experiments. Main features provided by
%Singularity are mobility of compute and reproducibility~\cite{kurtzer2017singularity}.
%We used Singularity version $2.5.2$ in our experiments.
%Users can easily access files, devices, sockets, ports of the host system from
%the Singularity container. By default, Singularity bind mounts /home/\$USER, /tmp, and \$PWD
%into the running container. Additional directories can be specified with \textit{--bind}
%to bind mount into the container. By defaut, Singularity uses as less as possible namespaces.
%However namespaces can be requested if required. Users cannot escalate privileges inside
%Singularty containers. This is because Singularity do not run containers as root.
%Instead, a user inside Singularity container is the same user outside the container.
%
%\subsection{Container Image Formats}
%
%A container image is a package that contains an application along with its
%required dependencies and libraries.
%Both, Docker and Singularity, have different image formats. Below we provide a description of both image
%formats.
%
%\subsubsection{Docker Image Format}
%
%Docker images are organized as a series of layers stacked on top of each
%other and identified by unique layer identifiers. If a layer is reused
%in different images it will
%have the same id across images. Each layer consists of a filesystem diff
%with respect to the layer below. 
%DockerHub uses gzip compression for
%transferring layers.There are two ways to create a Docker image. The
%first is to use an existing Docker image as a base image and then build a
%new image from that one, by making changes in the filesystem, which is
%saved as a new layer of the Docker image. The second way is to build a new
%Docker image from scratch.Docker supports several different storage drivers.
%\textit{Overlay2} is the preferred storage driver used by Docker for all currently
%supported Linux versions, whereas \textit{aufs} is preferred for Docker 18.06 and
%older, when using Ubuntu 14.04 on kernel 3.13 as it do not support \textit{overlay2}.
%Other storage drivers that are supported by Docker include \textit{devicemapper, btrfs,
%zfs, and vfs}
%
%\subsubsection{Singularity Image Format}
%
%On the other hand, a Singularity image is a single immutable file and 
%does not contain layers.
%Once a Singularity image is created, it cannot be changed, consequently,
%it can be used for reproducing the results of scientific
%experiments.
%Singularity takes snapshot, locks, and archives the developed application,
%which is known as Singularity image.
%It means that Singularity image is a single immutable file that can be moved around as
%any other file. This file can be checksummed, signed, and easily
%used for verification purposes. Singularity containers use Squashfs file format
%by default. It is read-only file system in Linux, which uses compression for files,
%inodes, and directories. However, Singularity also supports two other 
%formats i.e \textit{sandbox} format
%(which is just a chroot
%directory) and \textit{writable} format (the ext3 file system).
%While publishing experimental results, authors can also publish Singularity
%image along with its hash, which allow other researchers to verify results.
%\end{comment}
\subsection{Application frameworks}

We used containerized applications available in the Boutiques and BIDS Apps
frameworks used in neuroscience to run analyses on HPC clusters.

\subsubsection{Brain Imaging Data Structure (BIDS) Apps}

BIDS apps~\cite{gorgolewski2017bids} is a framework designed to share and execute neuroimaging
analysis pipelines. It improves usability, accesibility, and reproducibility
of different applications through containerization.
At the time of this study, there were 27 BIDS apps available and all are using different Docker images.
Out of these 27 images, we used 26 images for our experiments because one of them was not available
on the Docker hub.

\subsubsection{Boutiques}

Boutiques~\cite{glatard2018boutiques} is an application sharing system to
publish, integrate, and execute command-line applications across different
platforms. In Boutiques, the command-line of the application is described
through a flexible JSON descriptor pointing to a Docker or a Singularity
image where the application is installed. A set of core tools help
construct, validate, publish, and execute Boutiques descriptors. Boutiques
applications are published on the Zenodo research repository.

We used 49 Boutiques applications published on Zenodo at the time of
the writing. These applications use 23 different container images. Out of these images,
we could not get 3 images and 2 images were already included in BIDS images.
Therefore we are left with a total of 18 Boutiques images which include 6
Singularity images and 12 Docker images.

\begin{comment}
\subsubsection{Brain Imaging Data Structure (BIDS) Apps}

BIDS apps~\cite{gorgolewski2017bids} is a framework designed to share and execute neuroimaging
analysis pipelines. It improves usability, accesibility, and reproducibility
of different applications through containerization.
At the time of this study, there were 27 BIDS apps available and all are using different Docker images.
Out of these 27 images, we used 26 images for our experiments because the QAP bids app was not available
on Docker hub.
\end{comment}
So we have a total of 38 Docker images from both BIDS Apps and Boutiques.
	Table~\ref{table1} displays the names of all the images that we took in our experiments.
Table~\ref{sing} lists 6 Singularity images taken from the Boutiques framework.
%\begin{figure}
%\csvreader[%
% respect all,%
%  before reading=\footnotesize\sisetup{round-mode=places,round-precision=2,round-integer-to-decimal}
%      %\caption{Baseline GF Competition Circle Track 2014}\label{tab:baselline1}
%      \begin{adjustbox}{max width=\columnwidth},
%  after reading=\end{adjustbox},
% tabular={c |c | c | c | c},
%	table head =Image & {OS Distro} & {Vulnerabilities by Anchore} & {Vulnerabilities by Clair} & {Vulnerabilities by Vuls}\\\hline,
% late after line= \\,
%	late after last line=\\\hline %\multicolumn{5}{c}{Total Score: 1481}
%]{results.csv}{}{\csvlinetotablerow}%
%       \centering
%	\caption{Existing number of Vulnerabilities}
%\end{figure}

@@

\subsection{Scanning Images}

Scanning an image means looking into the image, collecting all the information about the image, and
finally, using this information to detect vulnerabilities present inside the image.
Scanning can be broadly categorized into two types: \rom{1}) Static Scanning and \rom{2})
Dynamic Scanning.
In Static scanning, an image is scanned without running it into a container. In respect to that,
different image features are extracted like installed packages, their versions, files, etc. to list
the vulnerabilities present inside the image.
On the other hand, Dynamic scanning involves running a container from the image and then scanning the
dynamic behavior of the image.

@@

We used three scanners namely Anchore~\cite{github_2019}, Vuls~\cite{future-architect_2019}, and
Clair Scanner~\cite{arminc_2019} for scanning Docker images because these are the most
mature scanners. Out of these scanners, only Vuls uses dynamic scanning and the remaining two
are static scanners.
Also, we froze vulnerability databases for these scanners on September 25, 2019, and scanned
all images to avoid getting different results due to the differences in the scanner's database.
Lastly, for scanning Singularity images, we used Stools.

@@

>>>>

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip}
\newcommand{\rom}[1]{\lowercase\expandafter{\romannumeral #1\relax}}
\newcommand{\tristan}[1]{\color{blue}From Tristan: #1\color{black}}

@@

\tristan{The above text contains true statements, but they are not organized 
to introduce the paper correctly. It should be rewritten according to the following logic:
\begin{itemize}
	\item Containers are popular in various scientific fields, in particular to run Big Data analyses on HPC clusters. Mention that neuroimaging is our particular focus, although this work is of general interest.
	\item However, deploying container images on clusters raises security issues. Some security issues that could possibly lead to privilege escalation, DoS or information leak. Some
	security issues are related to the architecture of container systems (Docker) but have been largely solved by Singularity.
	\item Here we are interested in image vulnerabilities [explain research questions as you did]
	\item [introduce paper outline as done]
\end{itemize}}

@@

This study involves four image scanning experiments.
The first experiment scans the images as is to report the existing vulnerabilities
inside the images. The scan report of
these images shows a large number of existing vulnerabilities which may pose
a potential security risk. These vulnerabilities exist due to the outdated
packages that are present inside the images. So, the next step is to
take the necessary steps to improve security in the images. Therefore, there exist
two possible solutions.
The first one involves updating
all these images and then rescanning them, which constitutes the second experiment of our study. The next 
experiment involves shrinking the size of the image
meaning that removing all unnecessary packages from the image to see the effect on the number of
vulnerabilities present inside the image, which will be presented as our third experiment.
Lastly, our fourth experiment involves both updating and shrinking the image size and then
rescanning for the vulnerabilities. \tristan{Content of this paragraph is fine, but writing is very convoluted.}

@@

\tristan{This sub-section has a lot of boilerplate material. It should be summarized.}

@@

%\begin{comment}
%Singularity offers mobility of compute by facilitating portable environments 
%through a single image file~\cite{kurtzer2017singularity}. Once Singularity image
%is created, it is hashed using SHA-256. Hence, it cannot be changed
%once it is created, consequently, it can be used for reproducing the results of
%scientific experiments. Main features provided by
%Singularity are mobility of compute and reproducibility~\cite{kurtzer2017singularity}.
%We used Singularity version $2.5.2$ in our experiments.
%Users can easily access files, devices, sockets, ports of the host system from
%the Singularity container. By default, Singularity bind mounts /home/\$USER, /tmp, and \$PWD
%into the running container. Additional directories can be specified with \textit{--bind}
%to bind mount into the container. By defaut, Singularity uses as less as possible namespaces.
%However namespaces can be requested if required. Users cannot escalate privileges inside
%Singularty containers. This is because Singularity do not run containers as root.
%Instead, a user inside Singularity container is the same user outside the container.
%
%\subsection{Container Image Formats}
%
%A container image is a package that contains an application along with its
%required dependencies and libraries.
%Both, Docker and Singularity, have different image formats. Below we provide a description of both image
%formats.
%
%\subsubsection{Docker Image Format}
%
%Docker images are organized as a series of layers stacked on top of each
%other and identified by unique layer identifiers. If a layer is reused
%in different images it will
%have the same id across images. Each layer consists of a filesystem diff
%with respect to the layer below. 
%DockerHub uses gzip compression for
%transferring layers.There are two ways to create a Docker image. The
%first is to use an existing Docker image as a base image and then build a
%new image from that one, by making changes in the filesystem, which is
%saved as a new layer of the Docker image. The second way is to build a new
%Docker image from scratch.Docker supports several different storage drivers.
%\textit{Overlay2} is the preferred storage driver used by Docker for all currently
%supported Linux versions, whereas \textit{aufs} is preferred for Docker 18.06 and
%older, when using Ubuntu 14.04 on kernel 3.13 as it do not support \textit{overlay2}.
%Other storage drivers that are supported by Docker include \textit{devicemapper, btrfs,
%zfs, and vfs}
%
%\subsubsection{Singularity Image Format}
%
%On the other hand, a Singularity image is a single immutable file and 
%does not contain layers.
%Once a Singularity image is created, it cannot be changed, consequently,
%it can be used for reproducing the results of scientific
%experiments.
%Singularity takes snapshot, locks, and archives the developed application,
%which is known as Singularity image.
%It means that Singularity image is a single immutable file that can be moved around as
%any other file. This file can be checksummed, signed, and easily
%used for verification purposes. Singularity containers use Squashfs file format
%by default. It is read-only file system in Linux, which uses compression for files,
%inodes, and directories. However, Singularity also supports two other 
%formats i.e \textit{sandbox} format
%(which is just a chroot
%directory) and \textit{writable} format (the ext3 file system).
%While publishing experimental results, authors can also publish Singularity
%image along with its hash, which allow other researchers to verify results.
%\end{comment}
\subsection{Container images}

We scanned a total of 44 container images taken from the BIDS
apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques}
frameworks, two containerization frameworks used in neuroscience to share
and re-use applications. Among the 27 BIDS apps available at the time of
the study \tristan{mention link where the apps are listed}, we studied 26
unique Docker images as one wasn't available on DockerHub. Among the 49
Boutiques applications available at the time of the study, listed using
Boutiques \texttt{bosh search} command, 23 unique container images were
listed including 3 that couldn't be retrieved and 2 that were already
included in BIDS apps. The resulting 18 images included 12 Docker images
and 6 Singularity images.
\tristan{I rewrote this paragraph to focus it on the information of relevance for the reader. 
You could take it as an example for other sections of the paper.}

%\begin{figure}
%\csvreader[%
% respect all,%
%  before reading=\footnotesize\sisetup{round-mode=places,round-precision=2,round-integer-to-decimal}
%      %\caption{Baseline GF Competition Circle Track 2014}\label{tab:baselline1}
%      \begin{adjustbox}{max width=\columnwidth},
%  after reading=\end{adjustbox},
% tabular={c |c | c | c | c},
%	table head =Image & {OS Distro} & {Vulnerabilities by Anchore} & {Vulnerabilities by Clair} & {Vulnerabilities by Vuls}\\\hline,
% late after line= \\,
%	late after last line=\\\hline %\multicolumn{5}{c}{Total Score: 1481}
%]{results.csv}{}{\csvlinetotablerow}%
%       \centering
%	\caption{Existing number of Vulnerabilities}
%\end{figure}

@@

\subsection{Image Scanners}

Container image scanners can be broadly categorized into static and dynamic
scanners. In static scanning, the image vulnerabilities are extracted from
file and package lists \tristan{Does it really look into files, or only installed packages? Please check. The difference is important 
when software is installed without using the package manager.}, without deploying the image in a container. In
dynamic scanning however, a container is started from the image and
\tristan{... It is still unclear how dynamic scanning works. Your text
mentions ``running a container from the image and then scanning the dynamic
behavior of the image'' but in practice it should explain what is executed
in the image: the package manager? other scanning tools? Please clarify by
completing this sentence.}

@@

We used four popular image scanners, namely Anchore~\cite{github_2019},
Vuls~\cite{future-architect_2019}, Clair Scanner~\cite{arminc_2019} and
Stools~\cite{refneeded}. Anchore, Vuls and Clair Scanner support Docker
images while Stools supports Singularity. Vuls is dynamic and the other
ones are static. To avoid discrepancies, we froze the vulnerability
databases for these scanners on September 25, 2019.

\tristan{The scanner descriptions need to be streamlined.}

@@

\tristan{Mention software not installed thorugh the package manager.}
